---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Load the activity data into the data frame variable : act. And correct the date format
```{r}
act <- read.csv("./05_ReproducibleResearch/week2/activity.csv")
act$date<- as.Date(act$date, format="%Y-%m-%d")
``` 
## What is mean total number of steps taken per day?
For this part of the assignment, we'll ignore the missing values in the dataset.
act1 : The subset of the activity data without missing steps values
```{r}
act1 <- subset(act, !is.na(steps))
``` 
calculate the sum steps by day, create  histogram of the daily total number of steps taken
```{r}
TotalStepsPerDay <- aggregate(act1$steps, by=list(date=act1$date), FUN=sum, na.rm = TRUE)
names(TotalStepsPerDay) <- c("date", "steps")
hist(TotalStepsPerDay$steps, breaks=5, xlab="Steps", main = "Total Steps per Day")
``` 

The mean and median total number of steps taken per day
mean(TotalStepsPerDay$steps, na.rm = TRUE)
median(TotalStepsPerDay$steps, na.rm = TRUE)

## What is the average daily activity pattern?
Calculate average steps for each interval for all days.
```{r}
AverageStepsPerInt <- aggregate(act1$steps, by=list(interval=act1$interval), FUN=mean, na.rm = TRUE)
names(AverageStepsPerInt) <- c("interval", "steps")
```
Plot the Average Number Steps per Day by Interval.

```{r}
with(AverageStepsPerInt,plot(interval,steps,type="l",xlab="5-minute intervals", ylab="Average number of steps across all days", col="red"))
```
interval with most average steps. 
```{r}
AverageStepsPerInt$interval[which.max(AverageStepsPerInt$steps)]
```
## Imputing missing values
the total number of missing values in the dataset 
```{r}
sum(is.na(act$steps))
```
Impute missing values strategy : Missing values will be imputed by inserting the average for each interval.

```{r}
act_NA <- act[is.na(act$step),]
act_merge <- merge(act_NA, AverageStepsPerInt, by=c("interval"))
act_merge1 <- act_merge[,c("steps.y","date","interval")]
names(act_merge1) <- names(act)
```
New dataset that is equal to the original dataset but with the missing data filled in :act2
```{r}
act2 <- rbind(act1, act_merge1)
```
calculate the sum steps by day, create  histogram of the daily total number of steps taken

```{r}
TotalStepsPerDay2 <- aggregate(act2$steps, by=list(date=act$date), FUN=sum, na.rm = TRUE)
names(TotalStepsPerDay2) <- c("date", "steps")
hist(TotalStepsPerDay2$steps, breaks=5, xlab="Steps", main = "Total Steps per Day")
```
The mean and median total number of steps taken per day

```{r}
mean(TotalStepsPerDay2$steps, na.rm = TRUE)
median(TotalStepsPerDay2$steps, na.rm = TRUE)
```

```{r}
```
## Are there differences in activity patterns between weekdays and weekends?
Create a new factor variable day_type in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day. 
```{r}
act2$day_type <- ifelse(weekdays(act2$date)=="Saturday" | weekdays(act2$date)=="Sunday", "Weekend", "Weekday")
```
Calculate average steps for each interval across all weekday days or weekend days.

```{r}
AverageStepsPerInt2 <- aggregate(act2$steps, by=list(act2$daytype, act2$interval), FUN=mean, na.rm = TRUE)
names(AverageStepsPerInt2) <- c("type", "interval","steps")
```
Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
library(lattice)

xyplot(steps ~ interval | daytype,
         layout = c(1, 2),
         xlab="Interval",
         ylab="Number of steps",
         type="l",
         lty=1,
         data=AverageStepsPerInt2)
```         
